---
title: "p8105_hw5_sd3730"
author: "Stacey Dai"
date: "2024-11-14"
output: github_document
---

```{r setup, message = FALSE, warning = FALSE}
library(ggplot2)
library(dplyr)
library(broom)
```

# Problem 1

First, I will create a function to check if there are duplicate birthdays in a group of 'n' people.
```{r dup_bd, message = FALSE, warning = FALSE}
has_duplicate_birthday = function(n) {
  birthdays = sample(1:365, n, replace = TRUE)
  
  return(any(duplicated(birthdays)))
}
```

Then, I will run the simulation 10,000 times for each group size and compute the probability of at least two people sharing a birthday.
```{r sim, message = FALSE, warning = FALSE}
set.seed(123)

probabilities = sapply(2:50, function(n) {
  mean(replicate(10000, has_duplicate_birthday(n)))
})
```

This is a plot that visualizes how the probability changes as the group size increases.
```{r bdayplot, message = FALSE, warning = FALSE}
prob_data = data.frame(
  group_size = 2:50, 
  probability = probabilities
)

ggplot(prob_data, aes(x = group_size, y = probability)) +
  geom_line(color = "blue") +
  labs(
    title = "Probability of Shared Birthday in a Group",
    x = "Group Size",
    y = "Probability of Shared Birthday"
  ) +
  theme_minimal()
```
The graph illustrate that as group size increases, the probability of at least two people sharing a birthday rises sharply. As the group size approaches 50, the probability nears 100%, meaning it's almost certain that two people will share a birthday in larger groups. This trend highlights how quickly the likelihood of shared birthdays increases with group size, even when birthdays are assumed to be evenly distributed.

# Problem 2

First, I will define the parameters.
```{r define, message = FALSE, warning = FALSE}
n = 30
sigma = 5
mu_vals = 0:6
alpha = 0.05
n_sim = 5000
```

Then, I will define a function to simulate data, calculate the sample mean and p-value, and return results for each dataset.
```{r ttestsim, message = FALSE, warning = FALSE}
simulate_t_test = function(mu, n, sigma, alpha) {
  sample_data = rnorm(n, mean = mu, sd = sigma)
  
  t_test_result = t.test(sample_data, mu = 0)
  
  test_summary = broom::tidy(t_test_result)
  
  tibble(
    mu_true = mu,
    mu_hat = test_summary$estimate,
    p_value = test_summary$p.value
  )
}
```

I'll run simulations for each true mean value.
```{r simtruemean, message = FALSE, warning = FALSE}
set.seed(123)

simulation_results = bind_rows(
  lapply(mu_vals, function(mu) {
    replicate(n_sim, simulate_t_test(mu, n, sigma, alpha), simplify = FALSE) |>
      bind_rows()
  })
)
```

